\thetitle

### 2021-05-15 paper/figures/productionsystem.tex
restructing

  \path (M1) edge[->, bend left=10, above]                 node[xshift=-0.1cm,yshift=-0.05cm] {$\scriptstyle 1,2,3$} (Q2);
  \path (M1) edge[->, bend right=10,below]                 node[xshift=-0.1cm,yshift=-0.05cm] {$\scriptstyle 4,5,6$} (Q3);
  \path (M2) edge[->, bend left=10,above,near start ]      node {$\scriptstyle 1$}     (Q4);
  \path (M2) edge[->, bend left=10,above,near start ]      node {$\scriptstyle 2$}     (Q5);
  \path (M2) edge[->, bend left=10,above,near start ]      node[xshift=0.07cm,yshift=-0.12cm] {$\scriptstyle 3$}     (Q6);
  \path (M3) edge[->, bend right=10,above,very near start] node[xshift=0.22cm,yshift=-0.09cm] {$\scriptstyle 4$}     (Q4);
  \path (M3) edge[->, bend right=10,above,very near start]      node[yshift=-0.075cm] {$\scriptstyle 5$}     (Q5);
  \path (M3) edge[->, bend right=10,above,near start]      node {$\scriptstyle 6$}     (Q6);
  \path (M4) edge[->, bend left=10,above,near start]       node[xshift=0.1cm] {$\scriptstyle 1,4$} (\fgi{});
  \path (M5) edge[->, bend left=10,above,near start]       node[xshift=0.075cm, yshift=-0.07cm] {$\scriptstyle 2,5$} (\fgi{});
  \path (M6) edge[->, bend right=10,below,near start]      node[xshift=0.15cm,yshift=0.07cm] {$\scriptstyle 3,6$} (\fgi{});

>>>>

  \path (M1) edge[->, bend left=10, above]                 node[xshift=-0.1cm,yshift=-0.05cm] {$\scriptstyle 1,2,3$} (Q2);
  \path (M1) edge[->, bend right=10,below]                 node[xshift=-0.1cm,yshift=-0.05cm] {$\scriptstyle 4,5,6$} (Q3);
  \path (M2) edge[->, bend left=10,above,near start ]      node {$\scriptstyle 1$}     (Q4);
  \path (M2) edge[->, bend left=10,above,near start ]      node {$\scriptstyle 2$}     (Q5);
  \path (M2) edge[->, bend left=10,above,near start ]      node[xshift=0.17cm,yshift=-0.2cm] {$\scriptstyle 3$}     (Q6);
960  \path (M3) edge[->, bend right=10,above,very near start] node[xshift=0.32cm,yshift=0.0cm] {$\scriptstyle 4$}     (Q4);
  \path (M3) edge[->, bend right=10,above,very near start]      node[yshift=-0.075cm] {$\scriptstyle 5$}     (Q5);
  \path (M3) edge[->, bend right=10,above,near start]      node {$\scriptstyle 6$}     (Q6);
  \path (M4) edge[->, bend left=10,above,near start]       node[xshift=0.1cm] {$\scriptstyle 1,4$} (\fgi{});
  \path (M5) edge[->, bend left=10,above,near start]       node[xshift=0.075cm, yshift=-0.07cm] {$\scriptstyle 2,5$} (\fgi{});
  \path (M6) edge[->, bend right=10,below,near start]      node[xshift=0.15cm,yshift=0.07cm] {$\scriptstyle 3,6$} (\fgi{});

### 2021-05-15 paper/paper.tex
restructing

\paragraph{Structure.} The rest of the paper is structured as follows. The next section introduces
average reward reinforcement learning and presents the used algorithm. For a more comprehensive
discussion of the theoretical foundations we refer to \citet{schneckenreither2020average}.
Section~\ref{sec:Experimental_Evaluation} describes the simulation model we use to evaluate the
approach. The results of the computation experiments are presented in
Section~\ref{sec:Computational}, and Section~\ref{sec:conclusion} concludes the paper.

@@

We hypothesis that dynamic lead time management approaches need to aim for a predictive design in
order to be able to compete with state-of-the-art order release methods from literature. However,
there only exist three papers that propose a predictive lead time management algorithm in
literature.

The first approach by \citet{PaterninaArboleda2001} introduces a predictive order release model by
using reinforcement learning in a single product, four-station serial flow line and compare its
performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP).
%
% Reinforcement learning is an optimisation technique that stems from dynamic programming and its goal
% is to find the best stationary policy for a given problem. This policy is usually provided by
% assessing current and future states (or state-action pairs) of an underlying Markov Decision Process
% (MDP).
% The advantage of reinforcement learning over dynamic programming is that (i) the problem
% space is explored by an agent and thus only expectantly interesting parts of the problem space need
% to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
% as the states are evaluated by consecutively observed states solely.
%
The algorithm of \citet{PaterninaArboleda2001} decides on whether or not to release an order after
each system change, the completion of an operation of any order at any stage or a new order arrival,
and assumes that any unsatisfied demand is lost. Thus, they use a continuous order release method
although in practice order release decisions often need to be made on a periodical basis, e.g.\@
daily (see \citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their tabular
based reinforcement learning agent.
%
Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
periodic order release decisions for a flow shop production system. The algorithm directly sets lead
times for each product type, which are then used to release the orders in the order pool. They show
that their approach outperforms static order release mechanisms by yielding lower costs, lateness
and standard deviation of lateness, but conclude that research using average reward reinforcement
learning methods harbor optimisation potential for the order release problem.
%
And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
times dynamically using an artificial neural network, which is used to forecast flow times. By
implementing a rolling horizon order release simulation of the proceeding periods they lift their
approach to a predictive lead time management approach, which is able to detect backorders of future
periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
prevent the negative effects of the lead time syndrome. Their model outperforms static and reactive
lead time management approaches, especially for scenarios with high utilisation.

\subsection{Related Reinforcement Learning Literature}

@@

\section{Average Reward Adjusted Reinforcement Learning}

This section briefly reintroduces the most important concepts of average reward adjusted
reinforcement learning, elaborates on optimality criteria and provides insights of the underlying
algorithm. For a more extensive introduction to average reward reinforcement learning we refer to
\citet{schneckenreither2020average}.

Like \cite{MillerVeinott1969} we are considering problems that are observed in a sequence of points
in time labeled \(1,2,\ldots\) and can be modelled using a finite set of states \(\States\),
labelled \(1,2,\ldots,\size{\States}\), where the size \(\size{\States}\) is the number of elements
in \(\States\). At each point $t$ in time the system is in a state \(s_{t} \in \States\). Further,
by choosing an action $a_{t}$ of a finite set of possible actions \(A_{s}\) the system returns a
reward $r_{t} = r(s_{t}, a_{t})$ and transitions to another state \(s_{t+1} \in \States\) at time
\(t+1\) with conditional probability \(p(s_{t+1}, r_{t} \mid s_{t}, a_{t})\). That is we assume that
reaching state \(s_{t+1}\) from state \(s_{t}\) with reward \(r_{t}\) depends solely on the previous
state \(s_{t}\) and chosen action \(a_{t}\). In other words, we expect the system to possess the
Markov property \citep[p.63]{sutton1998introduction}. Reinforcement learning processes that possess
the Markov property are referred to as Markov decision processes (MDPs)
\citep[p.66]{sutton1998introduction}.

@@

\paragraph{Average Reward Adjusted Deep RL Algorithm.} %
The model-free average reward adjusted RL algorithm is based on the tabular version of
\citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}. The algorithm operates
on a target network \(\netTarget{}\) and a worker network \(\netWorker{}\) as in
\citet{mnih2015human}. % The target network reduces fluctuations in the state value estimates and
% therefore reduces the
%
In the action selection process a \(\epsilon\)-sensitive lexicographic order
\(a=(a_{1},\ldots,a_{n}) \lex (b_{1},\ldots,b_{n})=b\) is used. It is defined as \(a \lex b\) if and
only if \(| a_{j} - b_{j} | \leqslant \epsilon\) for all \(j < i\) and
\(|a_{i} - b_{i}| > \epsilon\). Note that the resulting sets of actions may not be disjoint, but
taking the maximum as required in our algorithm is straight-forward and thus cheap to compute. The
first action selection criteria selects the actions which maximise the bias values, the second the
error term.
%
Equation~\ref{eq:constr2} of the average reward RL constraints is used to estimate the average
reward based on the currently inferred state values of two consecutive states and the returned
reward. % As the average
% reward is equal among all states estimating the average reward is based on many observations.
This infers a solid average reward prediction of the current policy in few steps only and thus
quickly leads to better policies, as ones with higher gain are preferred.
\citet{schneckenreither2020average} and
\citet{Tadepalli98_ModelbasedAverageRewardReinforcementLearning} find that updating the average
reward by Equation~\ref{eq:constr2} is superior as compared to exponentially smoothing the actual
observed rewards. This makes sense as the later rather evaluates past policies.

@@

This section introduces the simulation model and the parameterisation of the algorithm. To provide a
proof-of-concept for the proposed algorithm we adapt the flow shop presented in
\citet{schneckenreither2020order}. However, our algorithms operates on discrete lead times on a
product type basis and thus is not directly comparable to their order based lead time management
approach. Nonetheless, this section is largely based on the corresponding section of
\citet{schneckenreither2020order}.

@@

\paragraph{Demand.} The incoming orders are placed in the order pool with a due date slack of
\dds{} periods, that is each order arriving within period $t$ will be due at the end of period
$t+\dds{}$. The due date slack of \(\dds{}\) was chosen to (i) ensure sufficient time between the
first occurrence of orders in the order pool and the latest possible release of orders and (ii) to
have a clear cause and effect relationship between the order release decision and the cost
performance in the analysis.
%
To simulate medium and high variability of the demand process the interarrival time between
consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform ($\Unif$)
distribution, which are adjusted to yield the desired bottleneck utilisation level of either $70\%$
with $\Exp(135)$ and $\Unif(95,175)$, $80\%$ with $\Exp(118)$ and $\Unif(78,158)$, or $90\%$ with
$\Exp(105)$ and $\Unif(65,145)$. These values were chosen as they comprise the non-linear
relationship between flow times and high utilisation levels.

@@

@@

\subsection{Algorithm Setup}
\label{subsec:Algorithm_Setup}

@@

\paragraph*{Optimised Algorithm.}
Our implementation includes several optimisations, that help the algorithm in better exploring the
solution space and stabilise the learning process. These include the above discussed independent
agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous},
% overestimating the average reward \(\avgrew^{\pol}\)
and ensuring to constantly reduce the average reward as searching for policies with higher average
costs is unnecessary. Algorithm~\ref{alg:full}
depicts a detailed view on the algorithm including these extensions and with the objective to
minimise the returned rewards. The idea is the same as in Algorithm~\ref{alg:near}, except that the
action selection procedure is a loop over all independent agents and the implementation of the
n-step logic.
%
The algorithm is based around The agent first choose an action (steps \(4-5\)), executes the
selected action and stores the experience in the replay memory (step \(6\)) and then samples a
mini-batch of experiences each of length \(nstep\) to improve the policy (steps \(7-21\)). Step
\(17\) uses an exponentially smoothed overestimation (by \(2.5\%\)) of the average reward and rate
\(2^{-5}\). All average reward updates above this overestimation are rejected.

>>>>

\paragraph{Structure.} The rest of the paper is structured as follows. \MS[t]{TODO} The next section
introduces average reward reinforcement learning and presents the used algorithm. For a more
comprehensive discussion of the theoretical foundations we refer to
\citet{schneckenreither2020average}. Section~\ref{sec:Experimental_Evaluation} describes the
simulation model we use to evaluate the approach. The results of the computation experiments are
presented in Section~\ref{sec:Computational}, and Section~\ref{sec:conclusion} concludes the paper.

@@

\MS[]{todo: Reactive}

\MS[]{todo: proactive}

We hypothesise that dynamic lead time management approaches need to aim for a predictive design in
order to be able to compete with state-of-the-art order release methods from literature. However,
there only exist three papers that propose a predictive lead time management algorithm in
literature.

The first approach by \citet{PaterninaArboleda2001} introduces a predictive order release model by
using reinforcement learning in a single product, four-station serial flow line and compare its
performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP).
%
% Reinforcement learning is an optimisation technique that stems from dynamic programming and its goal
% is to find the best stationary policy for a given problem. This policy is usually provided by
% assessing current and future states (or state-action pairs) of an underlying Markov Decision Process
% (MDP).
% The advantage of reinforcement learning over dynamic programming is that (i) the problem
% space is explored by an agent and thus only expectantly interesting parts of the problem space need
% to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
% as the states are evaluated by consecutively observed states solely.
%
The algorithm of \citet{PaterninaArboleda2001} decides on whether or not to release an order after
each system change, the completion of an operation of any order at any stage or a new order arrival,
and assumes that any unsatisfied demand is lost. Thus, they use a continuous order release method
although in practice order release decisions often need to be made on a periodical basis, e.g.\@
daily (see \citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their tabular
based reinforcement learning agent.
%
Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
periodic order release decisions for a flow shop production system. The algorithm directly sets lead
times for each product type, which are then used to release the orders in the order pool. They show
that their approach outperforms static order release mechanisms by yielding lower costs, lateness
and standard deviation of lateness, but conclude that research using average reward reinforcement
learning methods harbor optimisation potential for the order release problem.
%
And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
times dynamically using an artificial neural network, which is used to forecast flow times. By
implementing a rolling horizon order release simulation of the proceeding periods they lift their
approach to a predictive lead time management approach, which is able to detect backorders of future
periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
prevent the negative effects of the syndrome. Their model outperforms static and reactive lead time
management approaches, especially for scenarios with high utilisation.

\subsection{Related Reinforcement Learning Applications}

@@

\subsection{Reinforcement Learning Preliminaries}

The rest of this section briefly reintroduces the important concepts of average reward adjusted
reinforcement learning, elaborates on optimality criteria and provides insights of the underlying
algorithm. For a more extensive introduction to average reward reinforcement learning we refer to
\citet{schneckenreither2020average}.

Like \citet{MillerVeinott1969} we are considering problems that are observed in a sequence of points
in time labeled \(1,2,\ldots\) and can be modelled using a finite set of states \(\States\),
labelled \(1,2,\ldots,\size{\States}\), where the size \(\size{\States}\) is the number of elements
in \(\States\). At each point $t$ in time the system is in a state \(s_{t} \in \States\). Further,
by choosing an action $a_{t}$ of a finite set of possible actions \(A_{s}\) the system returns a
reward $r_{t} = r(s_{t}, a_{t})$ and transitions to another state \(s_{t+1} \in \States\) at time
\(t+1\) with conditional probability \(p(s_{t+1}, r_{t} \mid s_{t}, a_{t})\). That is we assume that
reaching state \(s_{t+1}\) from state \(s_{t}\) with reward \(r_{t}\) depends solely on the previous
state \(s_{t}\) and chosen action \(a_{t}\). In other words, we expect the system to possess the
Markov property \citep[p.63]{sutton1998introduction}. Reinforcement learning processes that possess
the Markov property are referred to as Markov decision processes (MDPs)
\citep[p.66]{sutton1998introduction}.

@@

\section{Average Reward Adjusted Deep RL Algorithm }

The model-free average reward adjusted RL algorithm is based on the tabular version of
\citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}. The algorithm operates
on a target network \(\netTarget{}\) and a worker network \(\netWorker{}\) as in
\citet{mnih2015human}. % The target network reduces fluctuations in the state value estimates and
% therefore reduces the
%
In the action selection process a \(\epsilon\)-sensitive lexicographic order
\(a=(a_{1},\ldots,a_{n}) \lex (b_{1},\ldots,b_{n})=b\) is used. It is defined as \(a \lex b\) if and
only if \(| a_{j} - b_{j} | \leqslant \epsilon\) for all \(j < i\) and
\(|a_{i} - b_{i}| > \epsilon\). Note that the resulting sets of actions may not be disjoint, but
taking the maximum as required in our algorithm is straight-forward and thus cheap to compute. The
first action selection criteria selects the actions which maximise the bias values, the second the
error term.
%
Equation~\ref{eq:constr2} of the average reward RL constraints is used to estimate the average
reward based on the currently inferred state values of two consecutive states and the returned
reward. % As the average
% reward is equal among all states estimating the average reward is based on many observations.
This infers a solid average reward prediction of the current policy in few steps only and thus
quickly leads to better policies, as ones with higher gain are preferred.
\citet{schneckenreither2020average} and
\citet{Tadepalli98_ModelbasedAverageRewardReinforcementLearning} find that updating the average
reward by Equation~\ref{eq:constr2} is superior as compared to exponentially smoothing the actual
observed rewards. This makes sense as the later rather evaluates past policies.

@@

This section introduces the simulation model and the parameterisation of the algorithm. To provide a
proof-of-concept for the proposed algorithm we adapt the flow shop presented in
\citet{schneckenreither2020order}. However, our algorithms operates on discrete lead times on a
product type basis and thus is not directly comparable to their order based lead time management
approach. % Nonetheless, this section is largely based on the corresponding section of
% \citet{schneckenreither2020order}.

@@

\paragraph{Demand.} The incoming orders are placed in the order pool with a due date slack of
\dds{} periods, that is each order arriving within period $t$ will be due at the end of period
$t+\dds{}$. The due date slack of \(\dds{}\) was chosen to (i) ensure sufficient time between the
first occurrence of orders in the order pool and the latest possible release of orders and (ii) to
have a clear cause and effect relationship between the order release decision and the cost
performance in the analysis.
%
To simulate medium and high variability of the demand process the interarrival time between
consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform ($\Unif$)
distribution, which are adjusted to yield the desired bottleneck utilisation level of either $70\%$
with $\Exp(135)$ and $\Unif(95,175)$, $80\%$ with $\Exp(118)$ and $\Unif(78,158)$, or $90\%$ with
$\Exp(105)$ and $\Unif(65,145)$. These values were chosen as they comprise the non-linear
relationship between flow times and high utilisation levels.
%%%%%%
% Calculation
% (285 / (3 * u)) = interarrival time
%
% 285... Proc time bottleneck
% 3  ... every third product routes of this machine
% u  ... utilisation, e.g. 0.90

@@

\section{Lead Time Setting}

@@

\subsection{Average Reward Adjusted Deep RL Algorithm}
\label{subsec:Algorithm_Setup}

@@

\paragraph*{Optimised Algorithm.}
Our implementation includes several optimisations, that help the algorithm in better exploring the
solution space and stabilise the learning process. These include the above discussed independent
agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous},
% overestimating the average reward \(\avgrew^{\pol}\)
and ensuring to constantly reduce the average reward as searching for policies with higher average
costs is unnecessary. Algorithm~\ref{alg:full}
depicts a detailed view on the algorithm including these extensions and with the objective to
minimise the returned rewards. The idea is the same as in Algorithm~\ref{alg:near}, except that the
action selection procedure is a loop over all independent agents and the implementation of the
n-step logic.
%
The agent first choose an action (steps \(4-5\)), executes the selected action and stores the
experience in the replay memory (step \(6\)) and then samples a mini-batch of experiences each of
length \(nstep\) to improve the policy (steps \(7-21\)). Step \(17\) uses an exponentially smoothed
overestimation (by \(2.5\%\)) of the average reward and rate \(2^{-5}\). All average reward updates
above this overestimation are rejected.
